---
title: "CM Lab 2.4"
author: "Eric Stevens"
date: "10/16/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Environment Initialization
```{r}
library(fivethirtyeight) # new to you!
library(moderndive)
library(skimr)
library(tidyverse)
library(GGally) # new to you!
library(corrr)
```

## Data

The data we are interest is a subset of `hate_crimes` data set. The variables
we are interested in have the following meanings:

**avg_hatecrimes_per_100k_fbi:**
Average annual hate crimes per 100,000 population, FBI, 2010-2015

**gini_index:**
Gini Index, 2015

**share_pop_hs:**
Share of adults 25 and older with a high-school degree, 2009

**share_vote_trump:**
Share of 2016 U.S. presidential voters who voted for Donald Trump

## EDA Part 1

### Raw Data
```{r}
glimpse(hate_crimes)  
```

The data is made up of 12 variables, 11 numerical and 1 categorical, the state.
There is one row per state. 

```{r}
hate_crimes %>% 
  count(state, sort = TRUE)
```

From the table we can see that the data includes all 50 states and the 
District of Colombia.



### Summary Statistics

```{r}
crime <- select(hate_crimes, 
                    avg_hatecrimes_per_100k_fbi,
                    share_pop_hs,
                    gini_index,
                    share_vote_trump)

```


```{r}
skim(crime)
```



### Visualizations:
```{r}
ggplot(crime, aes(x = avg_hatecrimes_per_100k_fbi)) +
  geom_density() +
  labs(x = "", title = "Hate Crimes")
```

```{r}
# Histogram of share_pop_hs (IV):
ggplot(crime, aes(x = share_pop_hs)) +
  geom_density() +
  labs(x = "", title = "HS")
```

```{r}
# Histogram of gini (IV):
ggplot(crime, aes(x = gini_index)) +
  geom_density() +
  labs(x = "", title = "Gini")
```

```{r}
# Histogram of trump (IV):
ggplot(crime, aes(x = share_vote_trump)) +
  geom_density() +
  labs(x = "", title = "Trump")

```

```{r}
hate_demo <- crime %>% 
  mutate(
    cat_trump = case_when(
      share_vote_trump < .5 ~ "less than half", 
      TRUE ~ "more than half"
      )) %>% 
  mutate(cat_trump = as.factor(cat_trump)) %>% 
  select(-share_vote_trump)
```


## EDA Part 2
Now lets look at the way each variable correlates to the variable of interest.

```{r}
correlate(select(crime, avg_hatecrimes_per_100k_fbi, gini_index))
correlate(select(crime, avg_hatecrimes_per_100k_fbi, share_pop_hs))
correlate(select(crime, avg_hatecrimes_per_100k_fbi, share_vote_trump))
```

There is a 0.4212719 correlation between hate crimes and gini index value.
There is a 0.1641476 correlation between hate crimes and high school grad rate.
There is a 0.5045316 correlation between hate crimes and trump voter rate.

```{r}
ggplot(crime, aes(x = share_pop_hs, y = avg_hatecrimes_per_100k_fbi)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

ggplot(crime, aes(x = gini_index, y = avg_hatecrimes_per_100k_fbi)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

ggplot(crime, aes(x = share_vote_trump, y = avg_hatecrimes_per_100k_fbi)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

```

```{r}
ggpairs(hate_demo, aes(color = cat_trump))
```





## Multiple Regression Model

Now we will 

```{r}
hate_crime_model <- lm( avg_hatecrimes_per_100k_fbi ~ gini_index + share_pop_hs, data=crime)
get_regression_table(hate_crime_model)

```

```{r}
summary(hate_crime_model)
```

The regression table seems to indicate that when you combine the two 
independent variables into the model you end up with a very steep line. 
There is an intercept at -54 hate crimes, which is not possible, with a rise
of 64 hate crimes for a movement of one in the positive gini index axis and
a rise in 34 hate crimes for moving to 100 percent of the population HS 
educated.

This is drastically different than what we saw in the simple models, where 
there were not very steep lines cutting through the data. The multi-
regression seems to have caused a steep cut up through the data.


```{r}
library(plotly)
dim_scatter <- plot_ly(crime, 
                       x = ~share_pop_hs, 
                       y = ~gini_index, 
                       z = ~avg_hatecrimes_per_100k_fbi) %>%
  add_markers() %>%
  layout(scene = list(xaxis = list(title = 'HS'),
                     yaxis = list(title = 'Gini'),
                     zaxis = list(title = 'Hate Crimes')))
dim_scatter

```






