---
title: "MTH 530: Homework 4"
output: html_notebook
author: "Eric Stevens"
date: "November 29, 2018"
---

## Question 1

### a)

**Data:**
```{r}
inter_n <- 73
major_n <- 105
minor_n <- 240
no_n <- 1080


inter_xbar <- 6.22
major_xbar <- 5.81
minor_xbar <- 5.77
no_xbar <- 5.47


inter_sd <- 1.62
major_sd <- 1.43
minor_sd <- 1.24
no_sd <- 1.31
```

**Get Overall Mean**
In this case, the varying number of `n`s in each category results in the need to adjust the overall mean to be weighted by the number of `n`s. We can do this by multiplying our means in each category by their respective `n`, then sum these numbers, then divide by the total number of `n`s as follows.

```{r}
weighted_sum_of_means <- (inter_n*inter_xbar) +
  (major_n*major_xbar) +
  (minor_n*minor_xbar) +
  (no_n*no_xbar)

weighted_sum_of_means

sum_of_ns <- inter_n + major_n + minor_n + no_n

sum_of_ns

x_double_bar <- weighted_sum_of_means/sum_of_ns
x_double_bar

```

**Sum Of Squares**


* Between Groups
```{r}
inter_BT <- ((inter_xbar-x_double_bar)^2)*(inter_n)
major_BT <- ((major_xbar-x_double_bar)^2)*(major_n)
minor_BT <- ((minor_xbar-x_double_bar)^2)*(minor_n)
no_BT <- ((no_xbar-x_double_bar)^2)*(no_n)

BT <- inter_BT + major_BT + minor_BT + no_BT
BT

```


* Within
```{r}
inter_WI <- (inter_sd^2)*(inter_n-1)
major_WI <- (major_sd^2)*(major_n-1)
minor_WI <- (minor_sd^2)*(minor_n-1)
no_WI <- (no_sd^2)*(no_n-1)

WI <- inter_SSE + major_SSE + minor_SSE + no_SSE
WI
```

* Total
```{r}
SST = BT + WI
SST
```

**F Calculation**
```{r}
BToDF <- BT/(4-1)
WIoDF <- WI/(inter_n+major_n+minor_n+no_n-4)

F_value <- BToDF/WIoDF

BToDF
WIoDF
F_value

```



| -       | Sum of Squares | Degrees of Freedom | Mean Squared | F        |
|---------|----------------|--------------------|--------------|----------|
| Total   | 2677.968       | 1497               |              |          |
| Between | 57.18368       | 3                  | 19.06123     | 10.86601 |
| Within  | 2620.785       | 1494               | 1.754207     |          |






### b

**Get critical F value**
```{r}
qf(0.95, (4-1), inter_n+major_n+minor_n+no_n-4)
```
Since our F value is over 10 and our and our critical F value is 2.6, we can reject the null. This means that two or more of our categories are statistically significantly different from one and other. 

### c
* The initial data is normally distributed.
* Equal variance among groups.
* Samples are drawn independently of one and other.

### d
Yes, further analysis should be done to determine which mean or means have the statistically significant difference that is causing the F score allowing us to reject the null. 


## Question 2

```{r}
library(tidyverse)
library(resampledata)
cuckoos <- Cuckoos
head(cuckoos)
```

### a

```{r}
cuckoos %>% group_by(Bird) %>%
  ggplot(aes(x=Bird, y=Eggs)) +
  geom_boxplot()
```

### b
```{r}
# Compute the analysis of variance
the_anova <- aov(Eggs~Bird, data = cuckoos)
# Summary of the analysis
summary(the_anova)
```


### c

```{r}
TukeyHSD(the_anova) 
```


From this we can conclude that there is a statistically significant difference in egg size from each bird type to any other bird type.

### d

```{r}
TukeyHSD(the_anova, p.adjust(the_anova,method = "bonferroni"))
```

I cannot figure out how to do this. 