---
title: "Homework 3"
author: "Eric Stevens"
date: "November 13, 2018"
output: html_notebook
---


## Question 1

```{r}
bones <- data.frame(mother=1:10,
bf=c(1928, 2549, 2825, 1924, 1628, 2175, 2114, 2621, 1843, 2541),
pw=c(2126, 2885, 2895, 1942, 1750, 2184, 2164, 2626, 2006, 2627))
```


#### a) Null and alternative hypotheses:

$$H_0 :  \mu_{pw} - \mu_{bf} = 25$$
$$H_a : \mu_{pw} - \mu_{bf} > 25$$

#### b) The wrong test:

```{r}
# Remember, the two-sample test is inappropriate.
t.test(bones$pw, bones$bf, mu=25, alternative="greater", var.equal = TRUE)
```
Since $p>\alpha$ we cannot reject $H_0$, and therefore cannot conclude that 
it is highly unlikely that the observed statistics would occour if the $H_0$ 
was true.


#### c) The right test:
```{r}
# this is the right test
t.test(bones$pw, bones$bf, mu=25, paired = TRUE, alternative="greater")
```
Yes there is a signifigant difference in the result. Since $p<\alpha$, we 
can reject $H_0$, and conclude that it would be very unlikely to observe 
what we have if $H_0$ were true. The degree of freedom has been restricted 
compared to the two sample test because in this situation we are evaluating 
the mean difference in bone mass of individual mothers from brest feeding to
post weaning, where as in the two sample test we are evaluating the difference
in mean of the bresast feeding group as a whole and the post weaning group 
as a whole.

## Question 2

* a. She rejected her null hypothesis.
* b. The t-statistic based on her sample was greater than 2.01.
* c. The p value for her t-statistic was less than her α-level.
* e. Her sample mean of X was 9.664.


# Question 3

95% confident that the population mean falls between 0.99523 and 1.00077.


